---
title: "fazorsa_OriginalHomeworkCode_04"
author: "Frank Azorsa"
date: "10/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Brenna's Comments:
For Q1
Looking at the function, in the if statements, there was an error becasue you had 'SE.phat' labeled differently, with and without the ., so I added them in. 
I commented out the creation of NULL values, since they arent needed, and moved the instruction to print into the function
The math is the same in all, probably just to make testing the statements easier? I got all the equations from [Module 10](https://fuzzyatelin.github.io/bioanth-stats/module-10/module-10.html)
I also added a sample of the error messages I used for normality assumptions


***
###First question: 
Write a simple R function, Z.prop.test(), that can perform one- or two-sample Z-tests for proportion data, using the following guidelines:
Your function should take the following arguments: p1 and n1 (no default) representing the estimated proportion and sample size (i.e., based on your sample data); p2 and n2 (both defaulting to NULL) that contain a second sample’s proportion and sample size data in the event of a two-sample test; p0 (no default) as the expected value for the population proportion; and alternative (default “two.sided”) and conf.level (default 0.95), to be used in the same way as in the function t.test().
When conducting a two-sample test, it should be p1 that is tested as being smaller or larger than p2 when alternative=“less” or alternative=“greater”, the same as in the use of x and y in the function t.test().
The function should perform a one-sample Z-test using p1, n1, and p0 if either p2 or n2 (or both) is NULL.
The function should contain a check for the rules of thumb we have talked about (n∗p>5 and n∗(1−p)>5) to ensure the validity of assuming the normal distribution in both the one- and two-sample settings. If this is violated, the function should still complete but it should also print an appropriate warning message.
The function should return a list containing the members Z (the test statistic), P (the appropriate p value), and CI (the two-sided CI with respect to “conf.level” around p1 in the case of a one-sample test and around p2-p1 in the case of a two-sample test). For all test alternatives (“two.sided”, “greater”, “less”), calculate symmetric CIs based on quantiles of the normal distribution rather than worrying about calculating single-limit confidence bounds.
***
#im adding values= to start with to use in the function
```{r}
p1= 0.6
n1= 25
p0= 0.5
p2= 0.7
n2= 25
```

```{r}
z.prop.test <- function(p1, n1, p0, p2=NULL, n2=NULL, alternative="two.sided", conf.level=0.95) {
#   z <- NULL
#   ci <- NULL
#   p.val <- NULL
      phat <- p1/n1 #Sample proportion
   SE.phat <- NULL  
         if(alternative=="two.sided") {
         SE.phat <- sqrt(phat*(1-phat)/n1)# standar error of the sample proportion
         z <- (phat - p1)/SE.phat
         p.val <- pnorm(z, lower.tail = FALSE)# p-value for upper tail z-test
         lower <- phat - qnorm(conf.level)*SE.phat
         upper <- phat + qnorm(conf.level)*SE.phat
         ci <- c(lower, upper)         }
        
          if(alternative=="less") {
        SE.phat <- sqrt(phat*(1-phat)/n1)# standar error of the sample proportion
         z <- (phat - p1)/SE.phat
         p.val <- pnorm(z, lower.tail = FALSE)# p-value for upper tail z-test
         lower <- phat - qnorm(conf.level)*SE.phat
         upper <- phat + qnorm(conf.level)*SE.phat
         ci <- c(lower, upper)         }
         
         if(alternative=="greater") {
        SE.phat <- sqrt(phat*(1-phat)/n1)# standar error of the sample proportion
         z <- (phat - p1)/SE.phat
         p.val <- pnorm(z, lower.tail = FALSE)# p-value for upper tail z-test
         lower <- phat - qnorm(conf.level)*SE.phat
         upper <- phat + qnorm(conf.level)*SE.phat
         ci <- c(lower, upper)         }
    
         if(is.null(n2)|is.null(p2)) { 
         SE.phat <- sqrt(phat*(1-phat)/n1)# standar error of the sample proportion
         z <- (phat - p1)/SE.phat
         p.val <- pnorm(z, lower.tail = FALSE)# p-value for upper tail z-test
         lower <- phat - qnorm(conf.level)*SE.phat
         upper <- phat + qnorm(conf.level)*SE.phat
         ci <- c(lower, upper)
         } 
values <- c(z, p.val, ci)
print(values)
  if (n1*p0 < 5)
  {
  print("Caution, n*p in less than 5, violating normality assumption")
  }
   }
      #else {
         #SE.phat <- sqrt(phat*(1-phat)/n1)
      #}
      #ci <- phat + c( 
         #-1*((qnorm(((1 - conf.level)/2) + conf.level))*SE.phat),
         #((qnorm(((1 - conf.level)/2) + conf.level))*SE.phat))
   
      #return(list(estimate=phat, z=z, p.val=p.val, ci=ci))
      #}

```
***
###Challenges first question: I am still working on this function, and I am still getting errors and I am not sure if this is the correct way to present it - I will keep working.
***
***
###Second question:
The dataset from Kamilar and Cooper has in it a large number of variables related to life history and body size. For this exercise, the end aim is to fit a simple linear regression model to predict longevity (MaxLongevity_m) measured in months from species’ brain size (Brain_Size_Species_Mean) measured in grams. Do the following for both longevity~brain size and log(longevity)~log(brain size):
***
```{r}
#Loading the dataset “KamilarAndCooperData.csv” and the ggplot library
library(ggplot2)
library(curl)
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN597_Fall19/KamilarAndCooperData.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```
Fit the regression model and, using {ggplot2}, produce a scatterplot with the fitted line superimposed upon the data. Append the the fitted model equation to your plot (HINT: use the function geom_text()).

```{r}
lm_fit <- lm(MaxLongevity_m ~ Brain_Size_Species_Mean, data = d)
summary(lm_fit)
```
```{r}
predicted_d <- data.frame(MaxLongevity_m_pred = predict(lm_fit, d), Brain_Size_Species_Mean=d$Brain_Size_Species_Mean)
ggplot(data = d, aes(x = MaxLongevity_m, y = Brain_Size_Species_Mean)) + 
  geom_point(color='blue') +
  geom_line(color='red',data = predicted_d, aes(x=MaxLongevity_m_pred, y=Brain_Size_Species_Mean))
```

```{r}
ggplot(data = d, aes(x = MaxLongevity_m, y = Brain_Size_Species_Mean)) + 
  geom_point(color='blue') +
  geom_smooth(method = "lm", se = FALSE)
```
Identify and interpret the point estimate of the slope (β1), as well as the outcome of the test associated with the hypotheses H0: β1 = 0; HA: β1 ≠ 0. Also, find a 90 percent CI for the slope (β1) parameter.
```{r}
lm_fit
t <- coef(summary(lm_fit))
t <- data.frame(unlist(t))
colnames(t) <- c("Est", "SE", "t", "p")
t
```

```{r}
ci <- confint(lm_fit, level = 0.90)
ci
```
Using your model, add lines for the 90 percent confidence and prediction interval bands on the plot and add a legend to differentiate between the lines.

```{r}
ci <- predict(lm_fit, newdata = data.frame(Brain_Size_Species_Mean=d$Brain_Size_Species_Mean), interval = "confidence", 
    level = 0.90)
h_hat <- predict(lm_fit, newdata = data.frame(Brain_Size_Species_Mean=d$Brain_Size_Species_Mean))
df <- data.frame(cbind(d$Brain_Size_Species_Mean, d$MaxLongevity_m, h_hat))
names(df) <- c("x", "y", "yhat")
df <- cbind(df, ci)
names(df) <- c("x", "y", "yhat", "CIfit", "CIlwr", "CIupr")
head(df)
```
```{r}
g <- ggplot(data = df, aes(x = x, y = y))
g <- g + geom_point(alpha = 1/2)
g <- g + geom_line(aes(x = x, y = CIfit), colour = "black")
g <- g + geom_line(aes(x = x, y = CIlwr), colour = "blue")
g <- g + geom_line(aes(x = x, y = CIupr), colour = "blue")
g
```
Produce a point estimate and associated 90 percent PI for the longevity of a species whose brain weight is 800 gm. Do you trust the model to predict observations accurately for this value of the explanatory variable? Why or why not?

```{r}
ci <- predict(lm_fit, newdata = data.frame(Brain_Size_Species_Mean=800), interval = "confidence",
      level = 0.90)
ci
```
Looking at your two models, which do you think is better? Why?
```{r}

```
***
###Challenges second question: I still need to finish the last part of this question and test again each part of the code. 
***
